---
output:
  html_document: default
  pdf_document: default
---
Prosper Loan by Michael Roker
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(GGally)
library(dplyr)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data, message=FALSE, warning=FALSE}
# Load the Data
loandata <- read.csv('prosperLoanData.csv')

```



 This is an exploration of a dataset which approximately 113,937 records of personal loans taken from 
Prosper, a loan institution.

# Univariate Plots Section


```{r echo=FALSE, message=FALSE,  warning=FALSE, Univariate_Plots}
# look at the structure of the data and variables

str(loandata)

dim(loandata)

summary(loandata$IncomeRange)

```

As seen above, the dataset has 113,937 observations and 81 variables. All of 
these variables will not be explored going on this set.

```{r echo=FALSE, warning=FALSE, message=FALSE}

summary(loandata$DebtToIncomeRatio)

#Below shows the distribution of the debt to income ratio of the dataset.


# plot whole distribution
p1 = ggplot(data = subset(loandata, !is.na(DebtToIncomeRatio)), aes(DebtToIncomeRatio)) +
    geom_histogram(binwidth = 0.1, color = 'black', fill = '#099DD9') +
    ggtitle('Distribution of Debt To Income Ratio') +
    ylab('# of persons') +
    scale_y_log10()

# scale from 0 to 1.2
p2 = ggplot(data = subset(loandata, !is.na(DebtToIncomeRatio)), aes(DebtToIncomeRatio)) +
    geom_histogram(binwidth = 0.02, color = 'black', fill = '#099DD9') +
    scale_x_continuous(breaks = seq(0,1.2,.1), limits = c(0,1.2)) +
    ggtitle('Distribution of Debt To Income Ratio') +
    ylab('# of persons')

# scale from 0.7 to 10.1
p3 = ggplot(data = subset(loandata, !is.na(DebtToIncomeRatio)), aes(DebtToIncomeRatio)) +
    geom_histogram(binwidth = 0.1, color = 'black', fill = '#099DD9') +
    scale_x_continuous(breaks = seq(0.7,10.1,.5), limits = c(0.7,10.1)) +
    scale_y_continuous(breaks = seq(0,350,50), limits = c(0,350)) +
    ggtitle('Distribution of Debt To Income Ratio') +
    ylab('# of persons')

grid.arrange(p1,p2,p3)


```

Plotting the variable DebtToIncomeRatio, as shown above for the summary, 
the mean is 0.320 and the max is 10.010. Also, there are 8554 NA's which in
the plots have been omitted. The plots above give a better showing of
the summary which shows that the majority of the population are about 0.18. The 
first plot shows the whole distribution and the second and third zoom in on
0 - 1.2 and 0.7 - 10.1.

```{r echo=FALSE, warning=FALSE, message=FALSE}

summary(loandata$LoanStatus)

# plot two plots in grid for loan status
p1_loanstatus = ggplot(data = loandata, aes(LoanStatus)) +
  geom_histogram(stat = 'count')

# scale this plot log10
p2_loanstatus = ggplot(data = loandata, aes(LoanStatus)) +
  geom_histogram(stat = 'count') +
  scale_y_log10()

grid.arrange(p1_loanstatus,p2_loanstatus)

# plot for all of the past due dates
ggplot(data = subset(loandata, LoanStatus == "Past Due (16-30 days)" |
                       LoanStatus == "Past Due (>120)" |
                       LoanStatus == "Past Due (1-15 days)" |
                       LoanStatus == "Past Due (91-120 days)" |
                       LoanStatus == "Past Due (61-90 days)"),
       aes(LoanStatus)) +
  geom_histogram(stat = 'count')

```

This is the loan status variable. Firstly, looking at the different statuses 
and how they look in the population. Afterward, because the past due payments 
were so minimal, I zoomed in to see how they were compared to the other
statuses. Lastly, I took a view at only the past due payments variables to 
see how they compare with each other.


```{r echo=FALSE, warning=FALSE, message=FALSE}

summary(loandata$AmountDelinquent)

# show whole population
p1_amountdelinquent = ggplot(data = subset(loandata, !is.na(AmountDelinquent)), aes(AmountDelinquent)) +
    geom_histogram(color = 'orange') 
    #scale_x_continuous(limits = c(0,463881)) +
    #scale_y_continuous(limits = c(0,3500))

# show scale log 10
p2_amountdelinquent = ggplot(data = subset(loandata, !is.na(AmountDelinquent)), aes(AmountDelinquent)) +
    geom_histogram(color = 'orange') +
    scale_x_continuous(limits = c(0,463881)) +
    scale_y_continuous(limits = c(0,3500)) +
    scale_y_log10()

# scale for those over 1
p3_amountdelinquent = ggplot(data = subset(loandata, !is.na(AmountDelinquent)), aes(AmountDelinquent)) +
    geom_histogram(color = 'orange') +
    scale_x_continuous(limits = c(1,30000)) +
    scale_y_continuous(limits = c(0,1500))

grid.arrange(p1_amountdelinquent,p2_amountdelinquent, p3_amountdelinquent)



```

Exploring the Amount Delinquent variable, the mean shows at 984.5. There are
a number of NA's so those were omitted in the above plots. For the first
plot in the grid, it shows that the majority is nearer to 0 and that the max
is an outlier. For the second plot, the y axis was scaled up to take a 
closer look at the counts. For the third, the Amounts were scaled to 
remove the 0's (to look at those that were actually delinquent) and limited x
to see the distribution closer.


```{r echo=FALSE, warning=FALSE, message=FALSE}

summary(loandata$IncomeRange)

# plot all of the income ranges
ggplot(data = loandata, aes(IncomeRange)) +
  geom_bar(stat = 'count')


#remove not displayed, not employed, and 0 from incomerange to plot
incomerange_subset <- subset(loandata, IncomeRange != "Not displayed" & IncomeRange != "Not employed" & IncomeRange != "$0")


ggplot(data = incomerange_subset, aes(IncomeRange)) +
  geom_bar(stat = 'count')


```

As we can see for the summary, these are the amounts for the Income Range
variable. The first plot shows all of the income ranges. The second removes
not displayed, not employeyed, and 0.

```{r echo=FALSE, warning=FALSE, message=FALSE}

# remove 'other' from loandata dataset and plot
ggplot(data = subset(loandata, Occupation != 'other'), aes(Occupation)) + 
  geom_bar() +
  coord_flip() +
  scale_y_continuous(breaks = seq(0,5000,2000), limit = c(0,5000))


```

Among the top persons in the dataset are Teacher, Sales Executive, and Computer
Programmer.


```{r echo=FALSE, warning=FALSE, message=FALSE}

# showing summary of Listing Category using a factor
summary(factor(loandata$ListingCategory..numeric.))

# creating factored variable from numeric
loandata$ListingCategory..factor <- factor(loandata$ListingCategory..numeric.)

ggplot(data = loandata, aes(ListingCategory..factor, 
                            fill = factor(ListingCategory..factor,
                                           labels = c('0 Not Available', '1 Debt Consolidation',
                                                      '2 Home Improvement', '3 Business', '4 Personal Loan',
                                                      '5 Student Use', '6 Auto', '7 Other', '8 Baby & Adoption',
                                                      '9 Boat', '10 Cosmedic Procedure', '11 Engagement Ring',
                                                      '12 Green Loans', '13 Household Expenses', '14 large Purchases',
                                                      '15 Medical/Dental', '16 Motorcycle', '17 RV', '18 Taxes', '19 Vacation',
                                                      '20 Wedding Lons')))) + 
  geom_bar() +
  xlab('Listing Category')




# ggplot(data = loandata, aes(x = AmountDelinquent, color = ListingCategory..factor)) +
#   geom_freqpoly() +
#   scale_y_log10()
# 
# 
# ggplot(data = loandata, aes(x = ListingCategory..factor, y = AmountDelinquent)) +
#   geom_boxplot() +
# scale_y_continuous(limits = c(0,90000))



```

Changed this variable from numeric to category so that it can be used.
Created a bar plot to show the amounts for each category. As it can be seen 
Debt Consolidation is the dominant reason for getting a loan with Prosper.



```{r echo=FALSE, warning=FALSE, message=FALSE}
summary(loandata$LoanCurrentDaysDelinquent)


p1_LoanCurrentDaysDelinquent = ggplot(data = loandata, aes(LoanCurrentDaysDelinquent)) +
    geom_histogram(binwidth = 1) +
    scale_x_continuous(breaks = seq(0,2704,200), limits = c(0,2704)) +
    scale_y_continuous(breaks = seq(0,400,200), limits = c(0,200))


# scale by 10 and limit to 1000 days
p2_LoanCurrentDaysDelinquent = ggplot(data = loandata, aes(LoanCurrentDaysDelinquent)) +
    geom_histogram(binwidth = 10, color = 'black', fill = 'blue') +
    scale_x_continuous(breaks = seq(0,1000,100), limits = c(0,1000)) +
    scale_y_continuous(breaks = seq(0,400,200), limits = c(0,200)) +
    scale_y_log10()

grid.arrange(p1_LoanCurrentDaysDelinquent,p2_LoanCurrentDaysDelinquent)

```

As seen above for the summary of Days Delinquent for payment, the mean is 
152 days. The first plot is the whole population where 2704 is the max. The
second plot shows the plot scaled by 10 from 0 to 1000 days delinquent.



```{r echo=FALSE, warning=FALSE, message=FALSE}

summary(loandata$LoanOriginalAmount)

# look at whole spectrum of loans
p1_LoanOriginalAmount = ggplot(data = loandata, aes(LoanOriginalAmount)) +
    geom_histogram(color = 'black', fill = 'orange', binwidth = 1000) +
    scale_x_continuous(breaks = seq(1000,40000, 3000), limits = c(1000,40000))

# zoom into 25000 to 35000
p2_LoanOriginalAmount = ggplot(data = loandata, aes(LoanOriginalAmount)) +
    geom_histogram(color = 'black', fill = 'orange') +
    scale_x_continuous(breaks = seq(25000,35000, 1000), limits = c(25000,35000)) +
    scale_y_continuous(breaks = seq(0,1000,100), limits = c(0,1000))

# scale by 10
p3_LoanOriginalAmount = ggplot(data = loandata, aes(LoanOriginalAmount)) +
    geom_histogram(color = 'black', fill = 'orange', binwidth = 1000) +
    scale_x_continuous(breaks = seq(1000,40000, 3000), limits = c(1000,40000)) +
    scale_y_log10()

grid.arrange(p1_LoanOriginalAmount, p2_LoanOriginalAmount, p3_LoanOriginalAmount)

```

As shown in the summary above the mean for the amount Loaned is 8337. The
max borrowed is 35000. For all plots, changed the binwidth to 1000 because
loan amounts are measured by these amounts. In the third plot, scaled by 10. As
is shown above most persons borrowed 4000.

# Univariate Analysis

### Dataset Structure

This dataset has 113,937 observations along with 81 columns. The variables
being observed are:

* Debt to Income Ratio
* Loan Status
* Amount Delinquent
* Income Range
* Occupation
* Listing Category
* Loan Current Days Delinquent
* Loan Original Amount


### Features

The most interesting features in this dataset is the Loan Current Days 
Delinquent and the Loan Current Days Delinquent and if there is any 
commonalities between other variables and these two.

### Additional Features

I think that displaying the reasons for their loan, and possibly their
loan amounts may give understanding as to who mainly has problems
paying their loan, or if there is a trend.

### Changes made

The only change that was made was to the Listing Category, because this was in
numeric mode and this was changed to a factor so it can be displayed better
using a histogram. Also, so that it can be summarized with better clarity.


# Bivariate Plots Section


```{r echo=FALSE, warning=FALSE, message=FALSE}

cont_var <- select(loandata, LoanCurrentDaysDelinquent, LoanOriginalAmount, AmountDelinquent, DebtToIncomeRatio)

ggpairs(data = cont_var)

```


This gives a birds eye view of all of the different plots together along with 
some of the correlations.

```{r echo=FALSE, warning=FALSE, message=FALSE}

# filtering loandata to remove NA's and 0 from Amount Delinquent
loandata.filtered <- loandata %>% filter(!is.na(AmountDelinquent)) %>% filter(!is.na(DebtToIncomeRatio)) %>%
  filter(AmountDelinquent > 0)

cor.test(loandata.filtered$AmountDelinquent, loandata.filtered$DebtToIncomeRatio)


# plotting scatterplot of Debttoincomeratio vs Amountdelinquent, adding square root layer and summary line
ggplot(data = subset(loandata.filtered, AmountDelinquent > 0), aes(x = DebtToIncomeRatio, y = AmountDelinquent)) +
  geom_jitter(alpha = 6/20, color = 'orange') +
  scale_x_continuous(breaks = seq(0,1.0,0.1), limits = c(0,1.0)) +
  scale_y_continuous(limits = c(0,1e+05), breaks = seq(0,1e+05, 5000)) +
  geom_line(stat = 'summary', fun.y = mean, alpha = 3/10)


# this Amount Delinquent variable should not include 0, let's plot without that
ggplot(data = subset(loandata, AmountDelinquent > 0), aes(x = DebtToIncomeRatio, y = AmountDelinquent)) +
  geom_jitter(alpha = 3/20, color = 'blue') +
  xlim(0, quantile(loandata$DebtToIncomeRatio, 0.95,  na.rm = TRUE)) +
  ylim(0, quantile(loandata$AmountDelinquent, 0.95,  na.rm = TRUE)) +
  geom_smooth(method = 'lm', color = 'orange')



```

As seen above the correlation test is shown to be very low (-0.02) which
can indicate there is little pattern between the two variables.

For the first plot above, this was scaled between 0 and 1.0 seeing how 
the majority are between this amount. Added a summarized geom line on the plot.

On the second plot, used 95% quantile and added a geom smooth line to show
the relationship between Amount Delinquent and Debt To Income Ratio.

```{r echo=FALSE, warning=FALSE, message=FALSE, Bivariate_Plots}

cor.test(loandata$DebtToIncomeRatio, loandata$LoanCurrentDaysDelinquent)


# plot days delinquent vs debttoincomeratio 95% confidence interval 
ggplot(data = subset(loandata, LoanCurrentDaysDelinquent > 0), aes(x = DebtToIncomeRatio, y = LoanCurrentDaysDelinquent )) +
  geom_jitter(alpha = 4/20, color = 'orange') +
  ylim(0, quantile(loandata$LoanCurrentDaysDelinquent, 0.95, na.rm = TRUE)) +
  xlim(0, quantile(loandata$DebtToIncomeRatio, 0.95, na.rm = TRUE)) 



# for this function first is variable, second is categorial variable, third sum
# THIS IS FOR A DIFFERENT PLOT
by(loandata$AmountDelinquent, loandata$IncomeRange, summary)


# add variable to loandata df, to break up days into categories
#loandata$lcurrdysdel_breaks <- cut(loandata$LoanCurrentDaysDelinquent, breaks = seq(0,1500, 100))

#ggplot(data = loandata, aes(x = DebtToIncomeRatio, y = lcurrdysdel_breaks)) +
#  geom_point()


loandata.dir_round <- loandata

# round debto to income ratio to 1/10th.
loandata.dir_round$DebtToIncomeRatio = round(loandata.dir_round$DebtToIncomeRatio, digits = 1)



# additionaly for every debt to income ratio point make an average income
loandata.debtavg <- loandata.dir_round %>%
  filter(!is.na(LoanCurrentDaysDelinquent)) %>%
  group_by(DebtToIncomeRatio) %>%
  summarise(loancurrentdaysbreaks_mean = mean(LoanCurrentDaysDelinquent),
            n = n()) %>%
  arrange(DebtToIncomeRatio)

loandata.debtavg


# this plot shows for every debt to income ratio the average loan current days
ggplot(data =loandata.debtavg, aes(DebtToIncomeRatio, loancurrentdaysbreaks_mean)) +
  geom_line(alpha = 5/10)  +
  scale_x_continuous(breaks = seq(0,1.0,0.1), limits = c(0,1.0)) +
  scale_y_continuous(breaks = seq(0,500,100), limits = c(0,500)) +
  ggtitle('Average Loan Days Delinquent vs Debt To Income Ratio')

```

For the first plot, a 95% confidence interval was given, with a scatterplot
that was set to jitter.

For the second plot above, firstly, debttoincomeratio was rounded to 1/10's 
place. This would help with grouping the same. Secondly debttoincomeratio
was grouped by its values and the mean was taken for the loan days. This was 
then plotted with a line plot.


```{r echo=FALSE, warning=FALSE, message=FALSE}

cor.test(loandata$DebtToIncomeRatio, loandata$LoanOriginalAmount)


ggplot(data = loandata, aes(DebtToIncomeRatio, LoanOriginalAmount)) +
  geom_jitter(alpha = 4/20, color = 'blue') +
  scale_x_continuous(breaks = seq(0,5.0,0.5), limits = c(0,5.0)) +
  scale_y_continuous(breaks = seq(0,40000,5000))

ggplot(data = loandata, aes(DebtToIncomeRatio, LoanOriginalAmount)) +
  geom_jitter(alpha = 4/20, color = 'blue') +
  scale_x_continuous(breaks = seq(0,0.7,0.1), limits = c(0,0.7)) +
  scale_y_continuous(breaks = seq(0,40000,5000))
```
For the first plot, it can be seen that the majority of the population is up
to 1.0.


As seen above a plot was made comparing DebtToIncomeRatio to LoanOriginalAmount.
This was scaled for DebtToIncomeRatio to 0.7 being the maximum. As can be seen
there are amounts borrowed for every $5,000. However the trend can be seen
that as the ratio goes higher, the amounts borrowed goes lower.


```{r echo=FALSE, warning=FALSE, message=FALSE}

cor.test(loandata$AmountDelinquent, loandata$LoanCurrentDaysDelinquent)

ggplot(data = loandata, aes(LoanCurrentDaysDelinquent, AmountDelinquent)) +
  geom_point(alpha = 0.1) +
  scale_y_continuous(breaks = seq(0,10000,2000), limits = c(0,10000)) 


ggplot(data = subset(loandata, LoanCurrentDaysDelinquent > 0 & AmountDelinquent > 0), aes(LoanCurrentDaysDelinquent, AmountDelinquent)) +
  geom_point(alpha = 0.3) +
  scale_y_continuous(breaks = seq(0,10000,2000), limits = c(0,10000))


# create boxplot of days delinquent vs amountdelinquent
ggplot(data = subset(loandata, LoanCurrentDaysDelinquent > 0 & AmountDelinquent > 0), aes(LoanCurrentDaysDelinquent, AmountDelinquent)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq(0,10000,500), limits = c(0,10000))

```

The first plot above shows all values in the population. Alpha was reduced to
10% to show that the majority persons in this dataset is not delinquent
on both accounts.

For the second and third plot, the data was subsetted to omit 0 to take a closer
look at those who were delinquent (days and amounts).

# Bivariate Analysis

### Investigation

For Debt To Income Ratio vs Amount Delinquent, generally as the ratio got 
higher, the amount persons owed decreased. This was not a very 
strong correlation.

For Debt To Income Ratio vs the loan current days delinquent, 
this as well had no real pattern but was interesting to see that 
generally persons who had a higher debt ratio as expected they 
had more days delinquent in their payments.

For the debt to income ratio vs amount loaned, as expected the amounts 
borrowed became lower as the ratio was higher.

### Features

The most interesting feature found here was that as the ratio got higher
the Amount that a person was delinquent lowered. This was also the strongest
relationship found and most clear found.


# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots, warning=FALSE, message=FALSE}

# Select highest occupationloanamounts
loandata.Occ_red <- filter(loandata, Occupation %in% c("Commision", "Pharmacist", "Bus Driver", "Civil Service",
                                   "Engineer - Electrical", "Investor", "Flight Attendant",
                                   "College Freshman", "Tradesman - Carpenter", "Tradesman - Plumber"))

#loandata.Occ_red$DebtToIncomeRatio = round(loandata.Occ_red$DebtToIncomeRatio, digits = 1)



# look at for every debt to income ratio look at the average AmountDelinquent
# and split that by occupation

loandata.dir_round <- loandata.Occ_red %>%
  filter(!is.na(AmountDelinquent)) %>%
  group_by(DebtToIncomeRatio, Occupation) %>%
  summarise(amountdel_mean = mean(AmountDelinquent),
            n = n()) %>%
  ungroup() %>%
  arrange(DebtToIncomeRatio)

#loandata.dir_round$DebtToIncomeRatio = round(loandata.dir_round$DebtToIncomeRatio, digits = 1)




# look at the relationships between occupations
ggplot(data = loandata.Occ_red, aes(DebtToIncomeRatio, y = AmountDelinquent, color = Occupation)) +
  geom_jitter(alpha = 6/10) +
  scale_y_continuous(breaks = seq(0,100000,50000), limits = c(0,100000)) +
  scale_x_continuous(breaks = seq(0,0.7,0.1), limits = c(0,0.7)) +
  scale_y_log10()

# run a scatterplot for the average amount delinquent per occupation
ggplot(data = loandata.Occ_red, aes(x = DebtToIncomeRatio, y = AmountDelinquent, color = Occupation)) +
  geom_jitter() +
  scale_y_continuous(breaks = seq(0,5000, 500), limits = c(0,5000)) +
  scale_x_continuous(breaks = seq(0,1,.1), limits = c(0,1.0)) +
  scale_y_log10()

# create line plot to show averages accross occupations for amount delinquent
ggplot(data = loandata.dir_round, aes(x = DebtToIncomeRatio, y = amountdel_mean, color = Occupation)) +
  geom_line() +
  scale_x_continuous(breaks = seq(0,.7,.1), limits = c(0,.7)) 

# create square scale to see y axis closer
ggplot(data = loandata.dir_round, aes(x = DebtToIncomeRatio, y = amountdel_mean, color = Occupation)) +
  geom_line() +
  scale_x_continuous(breaks = seq(0,.7,.1), limits = c(0,.7)) +
  coord_trans(y = "sqrt") +
  scale_color_brewer()

```

Aboce, looking at the relationships between debt to income ratio and amount 
delinquent, no real relationship within Occupations. Above are the highest 
occupations in the for Amount Delinquent.

Looking at the different occupations in a geom line, we can see that the bus
drivers has some high spikes for amount delinquent.

```{r echo=FALSE, warning=FALSE, message=FALSE}

# find the highest reasons for taking out loans
by(loandata$AmountDelinquent, loandata$ListingCategory..factor, summary)

#reduce dataset to main highest loan reasons
loandata.ListingCat_red <- filter(loandata, ListingCategory..factor %in% c("1","6","10",
                                                                           "13","15","16"))
# get averages for debttoincomeratio and Listing category
groupings_debt_listing <- loandata.ListingCat_red %>%
  filter(!is.na(AmountDelinquent)) %>%
  group_by(DebtToIncomeRatio, ListingCategory..factor) %>%
  summarise(amountdel_mean = mean(AmountDelinquent),
            n = n()) %>%
  ungroup() %>%
  arrange(DebtToIncomeRatio)


ggplot(data = subset(groupings_debt_listing, amountdel_mean > 0), aes(x = amountdel_mean, y = DebtToIncomeRatio, color = ListingCategory..factor)) +
  geom_point() +
  scale_x_continuous(breaks = seq(0,1000,100), limits = c(0,1000)) +
  scale_y_continuous(breaks = seq(0,1.0,0.1), limits = c(0,1.0)) +
  scale_color_discrete(labels = c("1 - Debt Consolidation", "6 - Auto", "10 - Cosmetic Procedure",
                                             "13 - Household Expenses", "15 - Medical/Dental", "16 - Motorcycle"))


ggplot(data = groupings_debt_listing, aes(x = DebtToIncomeRatio, y = amountdel_mean, color = ListingCategory..factor)) +
  geom_line() +
  scale_x_continuous(breaks = seq(0,1.0,.1), limits = c(0,1.0)) +
  scale_color_discrete(labels = c("1 - Debt Consolidation", "6 - Auto", "10 - Cosmetic Procedure",
                                             "13 - Household Expenses", "15 - Medical/Dental", "16 - Motorcycle"))

# scale y axis to see closer
ggplot(data = groupings_debt_listing, aes(x = DebtToIncomeRatio, y = amountdel_mean, color = ListingCategory..factor)) +
  geom_line() +
  scale_x_continuous(breaks = seq(0,1.0,.1), limits = c(0,1.0)) +
  coord_trans(y = "sqrt") +
  scale_color_brewer(palette = 4, labels = c("1 - Debt Consolidation", "6 - Auto", "10 - Cosmetic Procedure",
                                             "13 - Household Expenses", "15 - Medical/Dental", "16 - Motorcycle"))

```

First, looked at all of the listing categories and the amount delinquent for
each one. Next, grouped all of the debt to income ratio and calculated the
means for each amount delinquent. This was done by summarising.

Created a scatter plot of all of the amounts accross the debt to income ratio.

The second plot created is a line plot which shows the same averages in 
comparison. 

Finally the third plot which is a line plot zooms in from 0 to 1.0 of 
debt to income ratio. 




# Multivariate Analysis

### Features

For the Amount Delinquent Mean plot, this would clearly show the different
Occupations and which ones have high means compared to
their initial debt to income ratio.

### Additional Features

The interesting factor is that cosmetic procedure seems to be a top factor
of persons who are in a delinquent status.


------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}

# plotting scatterplot of Debttoincomeratio vs Amountdelinquent, adding square root layer and summary line
ggplot(data = subset(loandata.filtered, AmountDelinquent > 0), aes(x = DebtToIncomeRatio, y = AmountDelinquent)) +
  geom_jitter(alpha = 6/20, color = 'orange') +
  scale_x_continuous(breaks = seq(0,1.0,0.1), limits = c(0,1.0)) +
  scale_y_continuous(limits = c(0,1e+05), breaks = seq(0,1e+05, 5000)) +
  geom_line(stat = 'summary', fun.y = mean, alpha = 3/10) +
  ggtitle('Amount Delinquent per Debt to Income Ratio')


```

This plot gives a good idea of the dataset and the fact that generally as 
the debttoincomeratio increases the amountdelinquent gradually decreases.
In other words, those who have good debt to income ratios will owe more, most
likely because they borrowed more in the first place. If those who have a 
really bad ratio owe less, its probably because they borrowed less.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}

# this plot shows for every debt to income ratio the average loan current days
ggplot(data =loandata.debtavg, aes(DebtToIncomeRatio, loancurrentdaysbreaks_mean)) +
  geom_line(alpha = 5/10)  +
  scale_x_continuous(breaks = seq(0,1.0,0.1), limits = c(0,1.0)) +
  scale_y_continuous(breaks = seq(0,500,100), limits = c(0,500)) +
  ggtitle('Average Loan Days Delinquent vs Debt To Income Ratio')

```

This plot showed the days that a person had a loan for based on their debt to
income ratio. This initially decreased but then gradually got higher as
their ratio went higher. This would show that those with higher debt to income
ratio has more days delinquent. This would make sense because those who have 
a worse reputatio more likely would have more days delinquent.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}

# create line plot to show averages accross occupations for amount delinquent
ggplot(data = loandata.dir_round, aes(x = DebtToIncomeRatio, y = amountdel_mean, color = Occupation)) +
  geom_line() +
  scale_x_continuous(breaks = seq(0,.7,.1), limits = c(0,.7)) +
  scale_y_continuous(breaks = seq(0,60000,10000)) +
  ggtitle('Amount Delinquent by Debt to Income Ratio by different Occupations') +
  ylab('Amount Delinquent Mean') +
  xlab('Debt To Income Ratio') 

```

This plot shows on the amount delinquent averages by the debt to income ratios
and this is separated by occupations. As we can see above, the tradesman
plumber has some high amounts owed as the ratio goes higher while Pharmacists
and Investors owe less as the ratio's are higher.
------

The Prosper Loan dataset contains 81 variables with 113,937 observations. I 
began by exploring the individual variables such as Debt to Income Ratio and
Amount Delinquent. A lot of reductions had to be made to variables like 
Occupation because there were many Occupations found within the dataset.

One trend found was Amount Delinquent and Debt To Income Ratio in that the 
higher the Debt To Income Ratio the less Amounts were Delinquent.
